---
title: "Formatting Data - annotated"
author: "Kaitlyn"
date: "7/17/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Moving the Formatting Data.R file into an RMarkdown document so that it is easier to annotate and dig into what is going on in the model.

This script loads the data and puts it into the appropriate format for Stan.

## Import and organize data

### Load detection matrices

```{r}
bobcat <- as.matrix(read.csv('Rota Data/Bobcat.csv', F))
coyote <- as.matrix(read.csv('Rota Data/Coyote.csv', F))
gryfox <- as.matrix(read.csv('Rota Data/Gray Fox.csv', F))
redfox <- as.matrix(read.csv('Rota Data/Red Fox.csv', F))

head(bobcat)
```

Looks like it's a detection matrix of 0 (non-detection), 1 (detection), NA (not operating). Each row is one camera site, and each column is a sampling period (ex. day).

### Calculate dates of camera operation

```{r}
cday <- apply(bobcat, 1, function(x) sum(is.na(x) == F))
head(cday)

```

Output is the number of days for which each camera was operating (0s or 1s; not NAs). These numbers are the same across species.

### Import covariates

```{r}
## covariates for probability of occurrence models
psi.cov <- read.csv('Rota Data/psi covariates.csv')
## covariates for detection probability models
p.cov <- read.csv('Rota Data/p covariates.csv')

head(psi.cov)
head(p.cov)
```

There is one row per camera, which has covariate values for occupancy (psi) and detection (p).

### Scale covariates for modeling

```{r}
d5km <- scale(psi.cov$Dist_5km)[, 1]
hden <- scale(psi.cov$HDens_5km)[, 1]
lati <- scale(psi.cov$Latitude)[, 1]
long <- scale(psi.cov$Longitude)[, 1]
lbyl <- lati * long
hike <- scale(psi.cov$People_site * 1000 / cday)
```

So the covariates are d5km (Dist_5km), hden (Hdens_5km), hike (People_site*1000/day), and then location (which has lati, longi, and lbyl which is a combination of latitude times longitude).

(Originally this code was lower down in the script but it makes sense logically for it to be here...)

## Create design matrix

### Create array

```{r}
X.array <- array(0, dim = c(nrow(bobcat), 16, 32))
```

This makes a blank matrix full of 0s. The x dimension correspond to the number of cameras (or, rows in bobcat). The y dimension corresponds to the unique combination of 0/1s across the 4 species (=16). The z dimension corresponds to the number of detection covariates (=32).

### Import general form of the design matrix

You need to specify the rows and columns that you include in the matrix to exclude the first two rows and the first columns (which include names for the columns/rows)
```{r}
library(xlsx) # install.packages("xlsx")
dm <- read.xlsx('Rota Data/Design Matrix.xlsx', 3, rowIndex = 3:18, colIndex = 2:33,
                header = F)
```

This gives a warning in the later versions of R, but it's okay. The 3 means it's taking from the third sheet of that excel file (where species interactions are influenced by covariates), and the row/colIndex tells which rows/columns to extract. header = F because the first element of the rowIndex vector does not contain the variables' names.

Template for the general form of the design matrix for the occupancy model linear predictor. Each row is a unique combination of latent species presence and absence (16 rows, 4x4 species) And each column represents a parameter. 

The rows are named 1111, 1110, 1101, 1100.... 0000; every possible combination of 1s and 0s. For reference, I will refer to this format as 1234 corresponding to the presence/absence of species 1, 2, 3, 4 (following the way Rota seems to do it).

A value of 1 indicates the parameter in the column is included in the linear predictor associated with the combination of latent presence and absence represented by the row; a value of 0 indicates the parameter is not included in the linear predictor for that particular combination of latent presence and absence. 

So, it looks like you fill this out ahead of time based on what covariates you hypothesize to influence different patterns of occupancy. There are 32 columns in the spreadsheet, nested under different "f" headings.

The first chunk seem like factors that influence the occupancy probability of a single species.
* F1 - 5 columns; have a value of "1" for all 8 rows in which species 1 is present
* F2 - 5 columns; have a value of "1" for all 8 rows in which species 2 is present
* F3 - 5 columns; have a value of "1" for all 8 rows in which species 3 is present
* F4 - 5 columns; have a value of "1" for all 8 rows in which species 4 is present

The next chunk seem like factors influencing the co-occurrence of pairs of species. Each pair of species has two columns because it doesn't assume the effect of species 1 on species 2 is the same as the effect of species 2 on species 1, for example. 

* F12 - 2 columns, have a value of "1" for all 4 rows in which species 1 and 2 are present
* F13 - 2 columns, have a value of "1" for all 4 rows in which species 1 and 3 are present
* F14 - 2 columns, have a value of "1" for all 4 rows in which species 1 and 4 are present
* F23 - 2 columns, have a value of "1" for all 4 rows in which species 2 and 3 are present
* F24 - 2 columns, have a value of "1" for all 4 rows in which species 2 and 4 are present
* F34 - 2 columns, have a value of "1" for all 4 rows in which species 3 and 4 are present

The values of the design matrix are filled in for each site in the next steps of this script.

### Fill in the design matrix with covariates of interest

```{r}
for(i in 1:nrow(bobcat)){ ## for every row (camera)
  for(j in 1:15){ # Maybe this is 15 rather than 16 because the last row is 0000 (so all species absent), and dm[j,] is never equal to 1
    X.array[i, j, dm[j, ] == 1] <- # the first dimension is the camera site; the second dimension is the covariate; the third is a series of true/false statements depending on whether that row... (type dm[1, ] == 1 for an example)
      c(1, lati[i], long[i], lbyl[i], hike[i], # individual species covariates
        1, lati[i], long[i], lbyl[i], hden[i], # individual species covariates
        1, lati[i], long[i], lbyl[i], hden[i], # individual species covariates
        1, lati[i], long[i], lbyl[i], hike[i], # individual species covariates
        1, d5km[i], # species interaction covariates
        1, hden[i], # species interaction covariates
        1, hden[i], # species interaction covariates
        1, d5km[i], # species interaction covariates
        1, hden[i], # species interaction covariates
        1, d5km[i])[dm[j, ] == 1]
  }
}
```

A little bit of detective work
```{r}
# what does this do?
dm[1, ] == 1 #when j = 1 (so the first species combo, 1234), returns TRUE/FALSE for whether a covariate is relevant
dm[2, ] == 1 #does the same for j = 2 (second species combo)  

# series of true false statements

length(dm[1, ] == 1) # 32 long; corresponds to a single row of the design matrix - whether or not it is equal to 1]

# this chunk of the for loop identifies the covariate values for a given camera i (in this case, i = 1, or the first camera)
c(1, lati[1], long[1], lbyl[1], hike[1], # individual species covariates
  1, lati[1], long[1], lbyl[1], hden[1], # individual species covariates
  1, lati[1], long[1], lbyl[1], hden[1], # individual species covariates
  1, lati[1], long[1], lbyl[1], hike[1], # individual species covariates
  1, d5km[1], # species interaction covariates
  1, hden[1], # species interaction covariates
  1, hden[1], # species interaction covariates
  1, d5km[1], # species interaction covariates
  1, hden[1], # species interaction covariates
  1, d5km[1])

# print one slice of matrix
head(X.array[,,1])
head(X.array[,1,])
head(X.array[1,,])

# now take a random slice too just for fun
head(X.array[,,5])
head(X.array[,5,])
head(X.array[5,,])
```


Seems like there are 5 covariates for each species. The first is just "1" so maybe this an intercept? Then latitude, longtiude, latitude*longitude, and a final covariate that differs for the four species (hikers for species 1 and 4, housing density for species 2 and 4). 

For the species co-occurrence covariates, there are 2 covariates. Again, the first is just "1" and the second varies by pairs. It's disturbed habitat for 1-2, 2-3, 3-4, and housing density for 1-3, 1-4, and 2-4.

This makes sense given the Rota et al text: 
Model M3 reﬂects the hypothesis that the relationship between human disturbance and occupancy probabilities for each species varies in the presence and absence of each of the other species. We modelled the probability that bobcats and coyote, coyote and grey fox, and grey fox and red fox occur together as a function of the proportion of recently disturbed habitat within 5 km; we modelled the probability that bobcat and grey fox, bobcat and red fox, and coyote and red fox occur together as a function of housing density within 5 km. There is little to guide decisions of what variables to use when modelling the probability that two or more of these species occur together, as this is one of the unique features of our model.

dm is the imported design matrix template with 0s and 1s, so it's only filling in the array when there is a "1" in this cell... ?

I still don't fully understand the X.array structure, but I think I have a good sense of what each of the lines of input code means, so I feel like I could manipulate to fit a different data structure.

### Determine whether or not species was observed at each site

```{r}
# OBSERVED Y
## function skips over any NA values
# the 1 indicates the function should be applied over the rows
y1max <- apply(bobcat, 1, max, na.rm = T)
y2max <- apply(coyote, 1, max, na.rm = T)
y3max <- apply(gryfox, 1, max, na.rm = T)
y4max <- apply(redfox, 1, max, na.rm = T)
```

The "max" values will either be 0 or 1—it's the maximum value of the detection matrix for a given species at a given site. So if 0, the species was never detected at that site; if 1, it was. This is used as input in the Stan model.

### Vectorizing detection/non-detection

This just efficiently makes a set of blank numeric vectors (Y1, Y2, Y3, Y4, trail, and dd) rather than doing it separately (ex. Y1 <- numeric() ; Y2 <- numeric() ; etc etc.)
```{r}
Y1 <- Y2 <- Y3 <- Y4 <- trail <- dd <- numeric()
```

Now we fill these vectors.

This loops through each of the cameras, and appends the values to the end of the vector (so it's not a matrix, but one vector with all of the values in it). 

As for what is in the values... For Y1, Y2, Y3, Y4, it's the detection record (0 or 1) for each day in which the camera was operating.

For trail, it is whether or not the camera was on trail (0 or 1) repeated for each day in which the camera was operating. For dd, it is the detection distance for the camera, again repeated for each day in which the camera was operating. rep means replicate. 
```{r}
for(i in 1:nrow(bobcat)){ 
  Y1 <- c(Y1, bobcat[i, 1:cday[i]]) 
  Y2 <- c(Y2, coyote[i, 1:cday[i]])
  Y3 <- c(Y3, gryfox[i, 1:cday[i]])
  Y4 <- c(Y4, redfox[i, 1:cday[i]])
  trail <- c(trail, rep(psi.cov$Trail[i], cday[i]))
  dd <- c(dd, rep(p.cov$Det_dist[i], cday[i]))
}
```

### Ragged indexing work-around?

Still not totally sure what this bit does:
Katie:
- nrow(bobcat) is the number of cameras
Is this equalizing the number of operational days somehow?
definition: "Ragged arrays are arrays that are not rectangular, but have different sized entries. This kind of structure crops up when there are different numbers of observations per entry."
```{r}
# Stan does not support ragged indexing.  This is a work-around
start <- 1
for(i in 2:nrow(bobcat)){
  start <- c(start, sum(cday[1:(i - 1)]) + 1)
}
```

### Define data (model input)

The descriptions of some of these values are in the model3.stan file, and others are to be found in the code above.

I have noted which ones are hard coded and need to be changed depending on our approach.
```{r}
# the descriptions of these values are in the model3.stan file
data <- list(
  K = ncol(X.array[1, , ]), # the number of occupancy parameters
  L = 3, # the number of detection parameters (is one of them an intercept..?) (HARD CODED)
  # their detection parameters were on/off trail and detection distance, second set of models also included presence/absence of coyotes for other species (maybe where the 3 is from?)
  N = nrow(bobcat), # the number of camera sites
  NJ = length(Y1), # the total number of observations (so the total number of camera days across all cameras)
  S = 16, # the number of unique combinations of 0s and 1s (across the 4 species) (HARD CODED)
  obs = cday, # from above, the number of days for which each camera was operating
  start = start, # 1, as defined above
  x = X.array, # matrix of covariate values
  trl = trail, # the trail detection covariate
  dd = scale(dd)[, 1], # the detection distance detection covariate (it is scaled in this line of code, as it is defined)
  I1 = y1max, I2 = y2max, I3 = y3max, I4 = y4max, # indications of whether or not species was ever detected at a given site (defined above; if 1, yes, if 0, no)
  Y1 = Y1, Y2 = Y2, Y3 = Y3, Y4 = Y4 # detection vectors
  )
```

### Define Bayesian priors (normal distributions)

```{r}
inits <- function(){
  list(
    a1 = rnorm(3),
    a2 = rnorm(3),
    a3 = rnorm(3),
    a4 = rnorm(3),
    beta = rnorm(ncol(X.array[1, , ])))
}
```

### Define model parameters

Initial guesses:
alpha 1, 2, 3, 4 = occupancy probabilities?
beta = covariate values?
ll = log likelihood?

```{r}
params <- c('a1', 'a2', 'a3', 'a4', 'beta', 'll')
```
