---
title: "GNP Models"
author: "Katie Grabowski"
date: "7/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
Here we go!
#install packages 
```{r}
library("unmarked")  #install.packages("unmarked") 
library(camtrapR) #install.packages("camtrapR")
library(tidyverse)
library(dplyr)
library(patchwork) #install.packages("patchwork")
library(ggpubr)  #install.packages("ggpubr")
```

#use unmarkedFrameOccuMulti to create the data object
```{r}
# load in detection histories
genet_dh <-read_csv("data/gorongosa-cameras/genet.csv", col_names = FALSE) %>% as.matrix()

civet_dh <-read_csv("data/gorongosa-cameras/civet.csv", col_names = FALSE) %>% as.matrix()

honeybadger_dh <-read_csv("data/gorongosa-cameras/honey_badger.csv", col_names = FALSE) %>% as.matrix()

marshmongoose_dh <-read_csv("data/gorongosa-cameras/marsh_mongoose.csv", col_names = FALSE) %>% as.matrix()

#I know I could have done this in the function, but separating it made more sense for me
y <- list(genet_dh, civet_dh, honeybadger_dh, marshmongoose_dh)
names(y) <- c('genet','civet','honey_badger', 'marsh_mongoose')

#trying to trouble shoot a below warning
any_na <- is.na(genet_dh) | is.na(civet_dh) | is.na(honeybadger_dh) | is.na(marshmongoose_dh) 
identical(genet_dh, civet_dh) #FALSE, which makes sense
identical(is.na(genet_dh), is.na(civet_dh)) #FALSE, which doesn't make sense
all.equal(is.na(genet_dh), is.na(civet_dh)) #1 element mismatch

#load occupancy covariates
#leaving the column names in for now
occ_covs <- read_csv("data/gorongosa-cameras/GNP covariates.csv", col_names = TRUE) %>% as.data.frame()

#scaling all the non-binary covariates to address the NaN warnings
occ_covs$urema_dist = scale(occ_covs$urema_dist)
occ_covs$tree_hansen = scale(occ_covs$tree_hansen)
occ_covs$termite.large.count.100m = scale(occ_covs$termite.large.count.100m)
occ_covs$lion_latedry = scale(occ_covs$lion_latedry)
occ_covs$cover.ground = scale(occ_covs$cover.ground)

# load in GNP cam metadata
cam_meta <- read_csv("data/gorongosa-cameras/cam_metadata_fromfield_and_raw_raster_withlion.csv")

# make p covariates table (detection)
GNP_p_covariates <- select(cam_meta, detect.obscured, cover.ground)

write_csv(GNP_p_covariates, "data/gorongosa-cameras/GNP p covariates.csv", col_names = T)

#this is a workaround because there's a bug in the package that doesn't allow me to put the detection covariates into siteCovs
#so this repeats each site-level detection covariate for every sample day
#only using this for the predict function
#July 2023 edit: this may be fixed by now, I'm not sure
det_covs <- GNP_p_covariates[rep(seq_len(nrow(GNP_p_covariates)), each = 122), ]

data_predict <- unmarkedFrameOccuMulti(y = y, siteCovs = occ_covs, obsCovs = det_covs)
#this line yields this warning: In unmarkedFrameOccuMulti(y = y, siteCovs = occ_covs, obsCovs = det_covs): Missing obs do not match among species. Converting some obs to NAs so they match.
#I think this means that the location of the NAs is not the same across all species, but I have no idea why that would be the case
#there are 459 NAs in civet, mm, and hb detection histories, and 458 in genet
#not sure where/why the difference is, but also not *too* worried about it?
#the warning suggests they just added an NA wherever that "extra" observation was

#same warning here, but I think it's fine
data <- unmarkedFrameOccuMulti(y = y, siteCovs = occ_covs)

#look at data
summary(data)
plot(data)

# Look at f parameter design matrix
data@fDesign
data@ylist
```
my f equation order: 
f1: genet
f2: civet
f3: honey badger
f4: marsh mongoose
f5: genet: civet
f6: genet: honey badger
f7: genet: marsh mongoose
f8: civet: honey badger
f9: civet: marsh mongoose
f10: honey badger: marsh mongoose

#run occuMulti
```{r}
# Formulas for state and detection processes
# dummy model

# Length should match number/order of columns in fDesign
occFormulas <- c('~urema_dist','~tree_hansen','~termite.large.count.100m','~lion_latedry','~1','~1','~1','~1','~1','~1', '0', '0', '0', '0', '0')

#Length should match number/order of species in data@ylist (genet, civet, honey badger, marsh mongoose)
detFormulas <- c('~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground')

fit <- occuMulti(detFormulas, stateformulas = occFormulas, data)
##^^yields a warning about missing detections that Ken Kellner said was okay to ignore
```

# explore model output
```{r}
#Look at output from dummy model
fit

#gives a list with one element for the state model
occ <- summary(fit[1]) 

write_csv(occ, "data/gorongosa-cameras/fit_occ.csv", col_names = T)

#plot(fit)

# occupancy only
fit@estimates[1]

# detection only
fit@estimates[2]

#leaving these in because they're potentially helpful to remember as functions that exist
unclass(fit) 
attr(fit, "estimates") 
as.data.frame(coef(fit))

#predict method
lapply(predict(fit,'state'),head)
lapply(predict(fit,'det'),head)

#marginal occupancy
predict(fit,'state',species='civet')
predict(fit,'state',species='genet') # why is this all 1 / 0 / 0?
predict(fit,'state',species='honey_badger')
predict(fit,'state',species='marsh_mongoose')
predict(fit,'det',  species='civet') # why are all of these the same?

#used to compare outcomes 
fl <- fitList(fit,fit2, fit3)
coef(fl)
```

#starting to think about actual models with ecological background [type 1, no species interactions]
```{r}
#null model for both occ and det
fit0 <- occuMulti(detformulas = c('~1', '~1', '~1', '~1'), stateformulas = c('~1', '~1', '~1', '~1','0','0','0','0','0','0','0','0','0','0','0'), data)

fit0

#null model for only occ
fit00 <- occuMulti(detformulas = c('~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground'), stateformulas = c('~1', '~1', '~1', '~1','0','0','0','0','0','0','0','0','0','0','0'), data)

fit00

# Length should match number/order of columns in fDesign
occFormulas1 <- c('~urema_dist','~urema_dist','~termite.large.count.100m','~urema_dist','0','0','0','0','0','0', '0', '0', '0', '0', '0')

#Length should match number/order of species in data@ylist (genet, civet, honey badger, marsh mongoose)
detFormulas <- c('~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground')

fit1 <- occuMulti(detFormulas, stateformulas = occFormulas1, data)

fit1

occFormulas2 <- c('~urema_dist+tree_hansen','~urema_dist + tree_hansen','~termite.large.count.100m + tree_hansen','~urema_dist + tree_hansen','0','0','0','0','0','0', '0', '0', '0', '0', '0')

fit2 <- occuMulti(detFormulas, stateformulas = occFormulas2, data)

fit2

#trying a model based on results from single species explorations
occFormulas3 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','0','0','0','0','0','0', '0', '0', '0', '0', '0')

fit3 <- occuMulti(detFormulas, stateformulas = occFormulas3, data)

fit3

#running fit3 twice, this time with the detection covariates in obsCovs so predict() runs
fit3_predict <- occuMulti(detFormulas, stateformulas = occFormulas3, data_predict)

#gives a list with one element for the state model
occ3 <- summary(fit3[1]) 
det3 <- summary(fit3[2])

write_csv(occ3, "data/gorongosa-cameras/occ3.csv", col_names = T)

#never writes row names
write_csv(det3, "data/gorongosa-cameras/det3.csv", col_names = T)

fit3

```

#Try to predict model output
```{r}
# explore "predict" function
predictions <- predict(fit3_predict, type = "state") # predicts based on the covariates at the 60 cameras
head(predictions$Predicted) # each row corresponds to camera, each column corresponds to a given state
rowSums(predictions$Predicted) # should sum to 1 (they do)
predictions$Predicted
predict(fit3_predict, type = "state", species = "civet")

# MAKE DUMMY DATA
# lake urema
urema_dist_scalelist <- list(scale = attr(occ_covs$urema_dist, "scaled:scale"),
                      center = attr(occ_covs$urema_dist, "scaled:center"))
dummy_urema <- data.frame(urema_dist = seq(min(occ_covs$urema_dist, na.rm=T), max(occ_covs$urema_dist, na.rm=T), by = 0.0001)) %>%
  mutate(termite.large.count.100m = 0,
         tree_hansen = 0,
         detect.obscured = 0,
         cover.ground = 0,
         urema_unscaled = (urema_dist * urema_dist_scalelist$scale + urema_dist_scalelist$center)/1000)

#tree_hansen
tree_hansen_scalelist <- list(scale = attr(occ_covs$tree_hansen, "scaled:scale"),
                      center = attr(occ_covs$tree_hansen, "scaled:center"))
dummy_tree <- data.frame(tree_hansen = seq(min(occ_covs$tree_hansen, na.rm=T), max(occ_covs$tree_hansen, na.rm=T), by = 0.0001)) %>%
  mutate(termite.large.count.100m = 0,
         urema_dist = 0,
         detect.obscured = 0,
         cover.ground = 0,
         tree_unscaled = (tree_hansen * tree_hansen_scalelist$scale + tree_hansen_scalelist$center))


#termite mound
termite_scalelist <- list(scale = attr(occ_covs$termite.large.count.100m, "scaled:scale"),
                      center = attr(occ_covs$termite.large.count.100m, "scaled:center"))
dummy_termite <- data.frame(termite.large.count.100m = seq(min(occ_covs$termite.large.count.100m, na.rm=T), max(occ_covs$termite.large.count.100m, na.rm=T), by = 0.0001)) %>%
  mutate(tree_hansen = 0,
         urema_dist = 0,
         detect.obscured = 0,
         cover.ground = 0,
         termite_unscaled = (termite.large.count.100m * termite_scalelist$scale + termite_scalelist$center))

#genet: black, 1
#civet: blue, 2
#honey badger: red, 3
#marsh mongoose: darkgreen, 4

# CIVET & LAKE UREMA 
civet_urema_predict <- predict(fit3_predict, type = 'state', newdata = dummy_urema, species="civet")
civet_urema_predict <- cbind(civet_urema_predict, dummy_urema)
civet_lake <- ggplot(data = civet_urema_predict, aes(x = urema_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "skyblue1") +
    geom_line(colour = "blue") +
    ylim(0,1) +
    theme_bw() +
    labs(x = "Distance to Lake (km)") +
    ggtitle("Civet")

#pdf("scripts/figures//civet_lake.pdf", width = 6, height = 5)
#civet_lake
#dev.off()

# GENET & LAKE UREMA (stays black/grey)
genet_urema_predict <- predict(fit3_predict, type = 'state', newdata = dummy_urema, species="genet")
genet_urema_predict <- cbind(genet_urema_predict, dummy_urema)
genet_lake <- ggplot(data = genet_urema_predict, aes(x = urema_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "navajowhite") +
    geom_line(colour = "darkorange") +
    ylim(0,1) +
    theme_bw() +
    labs(x = "Distance to Lake (km)") +
    ggtitle("Genet")

# HONEY BADGER & LAKE UREMA
honey_badger_urema_predict <- predict(fit3_predict, type = 'state', newdata = dummy_urema, species="honey_badger")
honey_badger_urema_predict <- cbind(honey_badger_urema_predict, dummy_urema)
honey_badger_lake <- ggplot(data = honey_badger_urema_predict, aes(x = urema_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "pink") +
    geom_line(colour = "red") +
    ylim(0,1) +
    theme_bw()+
    labs(x = "Distance to Lake (km)") +
    ggtitle("Honey Badger")

# MARSH MONGOOSE & LAKE UREMA
marsh_mongoose_urema_predict <- predict(fit3_predict, type = 'state', newdata = dummy_urema, species="marsh_mongoose")
marsh_mongoose_urema_predict <- cbind(marsh_mongoose_urema_predict, dummy_urema)
marsh_mongoose_lake <- ggplot(data = marsh_mongoose_urema_predict, aes(x = urema_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "darkseagreen2") +
    geom_line(colour = "darkgreen") +
    ylim(0,1) +
    theme_bw() +
    labs( x = "Distance to Lake (km)") +
    ggtitle("Marsh Mongoose")

#CIVET & TERMITE
civet_termite_predict <- predict(fit3_predict, type = 'state', newdata = dummy_termite, species="civet")
civet_termite_predict <- cbind(civet_termite_predict, dummy_termite)
civet_termite <- ggplot(data = civet_termite_predict, aes(x = termite_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "skyblue1") +
    geom_line(colour = "blue") +
    ylim(0,1) +
    theme_bw() +
    labs(x = "Termite Mound Density") +
    ggtitle("")

#MARSH MONGOOSE & TERMITE
marsh_mongoose_termite_predict <- predict(fit3_predict, type = 'state', newdata = dummy_termite, species="marsh_mongoose")
marsh_mongoose_termite_predict <- cbind(marsh_mongoose_termite_predict, dummy_termite)
marsh_mongoose_termite <- ggplot(data = marsh_mongoose_termite_predict, aes(x = termite_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "darkseagreen2") +
    geom_line(colour = "darkgreen") +
    ylim(0,1) +
    theme_bw() +
    labs( x = "Termite Mound Density") +
    ggtitle("Marsh Mongoose")

#HONEY BADGER & TREE
honey_badger_tree_predict <- predict(fit3_predict, type = 'state', newdata = dummy_tree, species="honey_badger")
honey_badger_tree_predict <- cbind(honey_badger_tree_predict, dummy_tree)
honey_badger_tree <- ggplot(data = honey_badger_tree_predict, aes(x = tree_unscaled, y = Predicted)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "pink") +
    geom_line(colour = "red") +
    ylim(0,1) +
    theme_bw() +
    labs( x = "Tree Cover (%)") +
    ggtitle("")


# now put everything together with patchwork!
#grid <- (civet_lake + civet_termite) /
#        (honey_badger_lake + honey_badger_tree) /
#        (genet_lake + marsh_mongoose_termite)
  # add plot labels
# plot_annotation(title = "Marginal Occupancies", tag_suffix = ")")

#create and save figure with all components
pdf("scripts/figures/best_interaction_model_predict.pdf", width = 6, height = 5)
ggarrange(civet_lake, civet_termite, honey_badger_lake, honey_badger_tree, genet_lake, marsh_mongoose_termite, ncol = 2, nrow = 3)
dev.off()
```

#starting to think about actual models with ecological background [type 2, no interaction covariates]
```{r}
#null model
fit01 <- occuMulti(detformulas = c('~1', '~1', '~1', '~1'), stateformulas = c('~1', '~1', '~1', '~1','~1','~1','~1','~1','~1','~1','0','0','0','0','0'), data)

fit01

#null model
fit001 <- occuMulti(detformulas = c('~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground'), stateformulas = c('~1', '~1', '~1', '~1','~1','~1','~1','~1','~1','~1','0','0','0','0','0'), data)

fit001

# Length should match number/order of columns in fDesign
#single 'best' occ covariate from reading
occFormulas11 <- c('~urema_dist','~urema_dist','~termite.large.count.100m','~urema_dist','~1','~1','~1','~1','~1','~1', '0', '0', '0', '0', '0')

#Length should match number/order of species in data@ylist (genet, civet, honey badger, marsh mongoose)
detFormulas <- c('~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground', '~detect.obscured+cover.ground')

fit11 <- occuMulti(detFormulas, stateformulas = occFormulas11, data)

#why does this not produce anything except estimates?
fit11

#top two occ covariates from reading
occFormulas12 <- c('~urema_dist+tree_hansen','~urema_dist + tree_hansen','~termite.large.count.100m + tree_hansen','~urema_dist + tree_hansen','~1','~1','~1','~1','~1','~1', '0', '0', '0', '0', '0')

fit12 <- occuMulti(detFormulas, stateformulas = occFormulas12, data)

fit12

#best occ covariates from single species exploration
occFormulas13 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~1','~1','~1','~1','~1','~1', '0', '0', '0', '0', '0')

fit13 <- occuMulti(detFormulas, stateformulas = occFormulas13, data)

fit13

```

#starting to think about actual models with ecological background [type 3, includes interaction covariates] 

```{r}
# Length should match number/order of columns in fDesign
occFormulas21 <- c('~urema_dist','~urema_dist','~termite.large.count.100m','~urema_dist','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry', '0', '0', '0', '0', '0')

fit21 <- occuMulti(detFormulas, stateformulas = occFormulas21, data)

fit21

#top two occ covariates from reading
occFormulas22 <- c('~urema_dist+tree_hansen','~urema_dist + tree_hansen','~termite.large.count.100m + tree_hansen','~urema_dist + tree_hansen','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry', '0', '0', '0', '0', '0')

fit22 <- occuMulti(detFormulas, stateformulas = occFormulas22, data)

fit22

#best occ covariates from single species exploration
occFormulas23 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry','~lion_latedry', '0', '0', '0', '0', '0')

fit23 <- occuMulti(detFormulas, stateformulas = occFormulas23, data)

fit23
```

#trying to see when the model breaks with the interaction term
```{r}
#starting with only an interaction covariate for genet/civet
occFormulas31 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~1','~1','~1','~1','~1', '0', '0', '0', '0', '0')

fit31 <- occuMulti(detFormulas, stateformulas = occFormulas31, data)

fit31

occFormulas32 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~lion_latedry','~1','~1','~1','~1', '0', '0', '0', '0', '0')

fit32 <- occuMulti(detFormulas, stateformulas = occFormulas32, data)

fit32

occFormulas33 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~1','~lion_latedry','~1','~1','~1', '0', '0', '0', '0', '0')

fit33 <- occuMulti(detFormulas, stateformulas = occFormulas33, data)

fit33

#34 didn't even run...
occFormulas34 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~1','~1','~lion_latedry','~1','~1', '0', '0', '0', '0', '0')

fit34 <- occuMulti(detFormulas, stateformulas = occFormulas34, data)

fit34

occFormulas35 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~1','~1','~1','~lion_latedry','~1', '0', '0', '0', '0', '0')

fit35 <- occuMulti(detFormulas, stateformulas = occFormulas35, data)

fit35

occFormulas36 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~lion_latedry','~1','~1','~1','~1','~lion_latedry', '0', '0', '0', '0', '0')

fit36 <- occuMulti(detFormulas, stateformulas = occFormulas36, data)

fit36

occFormulas37 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~1','~lion_latedry','~1','~1','~1','~1', '0', '0', '0', '0', '0')

fit37 <- occuMulti(detFormulas, stateformulas = occFormulas37, data)

fit37

occFormulas38 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~1','~1','~lion_latedry','~1','~1','~1', '0', '0', '0', '0', '0')

fit38 <- occuMulti(detFormulas, stateformulas = occFormulas38, data)

fit38

occFormulas39 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~1','~1','~1','~lion_latedry','~1','~1', '0', '0', '0', '0', '0')

fit39 <- occuMulti(detFormulas, stateformulas = occFormulas39, data)

fit39

occFormulas40 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~1','~1','~1','~1','~lion_latedry','~1', '0', '0', '0', '0', '0')

fit40 <- occuMulti(detFormulas, stateformulas = occFormulas40, data)

fit40

occFormulas41 <- c('~urema_dist','~urema_dist + termite.large.count.100m','~urema_dist + tree_hansen','~termite.large.count.100m','~1','~1','~1','~1','~1','~lion_latedry', '0', '0', '0', '0', '0')

fit41 <- occuMulti(detFormulas, stateformulas = occFormulas41, data)

fit41


```
#model comparisons
```{r}
#List of fitted models
fl <- fitList(fit0, fit00, fit1, fit2, fit3, fit01, fit001, fit11, fit12, fit13, fit21, fit22, fit23)
coef(fl)

#Model selection
modsel <- modSel(fl)
modSel(fl)
summary(modsel)

ms <- modSel(fl, nullmod="fit0")
ms
 
coef(ms)                            # Estimates only
SE(ms)                              # Standard errors only
(toExport <- as(ms, "data.frame"))  # Everything as a data frame

```

```{r}
#conditional occupancy
head(predict(fit3,'state',species='civet')) #civet | genet present
head(predict(fit,'state',species='tiger',cond='bear')) #tiger | bear present
head(predict(fit,'state',species='tiger',cond='-bear')) #bear absent
head(predict(fit,'state',species='tiger',cond=c('coyote','-bear')))

```

