---
title: "Supplementary Information: Simulation Results"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, echo=FALSE, warning=FALSE, fig.height=7, fig.width=9)
```

```{r}
library(tidyverse)
source('sim_functions.R')

f <- c(1.7,-0.7,1)
psi <- c(exp(sum(f)),exp(f[1]),exp(f[2]),exp(0))
psi <- psi/sum(psi)
#psi

N <- 100

#True occupancy probability for each state
#psi <- c(0.5, 0.4, 0.03, 0.07)

#Probability of initial markov state
mu1 <- c(3, 150)
mu2 <- c(1, 60)

#coyotes: 0.68, 0.34, 0.17
#deer: 1, 0.5, 0.25

#Detection intensities (always 0 in first state)
#lambdas are rough rounding of values from fitted model
#can't set first value to 0, has to at least be a little bigger
lam1 <- c(0.0001,40) #sp1 | sp2 present
lam2 <- c(0.0001,11) #sp1 | sp2 absent
lam3 <- c(0.0001,3.5) #sp2

#Time intervals
maxt <- 21 #3 days, intervals by hour


sim_truth <- c(f, log(mu1), log(mu2), log(lam1[2]), log(lam2[2]), log(lam3[2]))

# Find values for mu[1]

#Deer
mu_deer_vals <- rep(NA,3)
cands <- as.matrix(data.frame(m1=seq(0.05,3,0.001), m2=mu1[2]))
pval <- rep(NA, nrow(cands))
for (i in 1:length(pval)){
  pval[i] <- get_p(cands[i,], lam1, maxt)
}
mu_deer_vals[1] <- cands[,1][which(round(pval,2)==0.25)[1]]
mu_deer_vals[2] <- cands[,1][which(round(pval,2)==0.5)[1]]
mu_deer_vals[3] <- mu1[1]

#plot(cands[,1], pval, type='l')
#for (i in 1:3){
#  abline(v=mu_deer_vals[i], col='red')
#}

#Coyote
mu_coy_vals <- rep(NA,3)
cands <- as.matrix(data.frame(m1=seq(0.05,1,0.001), m2=mu2[2]))
pval <- rep(NA, nrow(cands))
for (i in 1:length(pval)){
  pval[i] <- get_p(cands[i,], lam3, maxt)
}
mu_coy_vals[1] <- cands[,1][which(round(pval,2)==0.25)[1]]
mu_coy_vals[2] <- cands[,1][which(round(pval,2)==0.5)[1]]
mu_coy_vals[3] <- mu2[1]

#plot(cands[,1], pval, type='l')
#for (i in 1:3){
#  abline(v=mu_coy_vals[i], col='red')
#}

# Set up simulation matrix

Nsize <- c(100,1000)

sim_mat <- expand.grid(mu_deer_vals, mu_coy_vals, Nsize)

colnames(sim_mat) <- c("mu1_1","mu1_2","N")
sim_mat$ID <- 1:18
sim_mat$sp1 <- c("low","med","high")
sim_mat$sp2 <- rep(c("low","med","high"),each=3)


# Load in and format results

truth <- c(psi, sim_truth[4:10])

sim_N100 <- readRDS('sim_N100.Rds')

sim_N1000 <- readRDS('sim_N1000.Rds')

pars <- c("both","deer","coyote","none","mu1[1]","mu1[2]","mu2[1]","mu2[2]","lam1","lam2","lam3",
          "p1_2pres","p1_2abs","p2","sp1_dur1","sp1_dur2","sp2_dur1","sp2_dur2")

sim_df <- lapply(1:9, function(i){
  sublist <- sim_N100[[i]]
  nsim <- length(sublist)
  out <- lapply(1:nsim, function(x){
    if(is.null(sublist[[x]])) return(data.frame(ID=i, repl=x, param=pars, val=NA,truth=NA,conv=1,
                                     lik=NA))
    est <- sublist[[x]]$par
    psi <- c(exp(est[1]+est[2]+est[3]),exp(est[1]),exp(est[2]),exp(0))
    psi <- psi/sum(psi)
    est <- c(psi, exp(est[4:10]))
    new_truth <- truth
    new_truth[5:11] <- exp(new_truth[5:11])
    new_truth[5] <- sim_mat[i,1]
    new_truth[7] <- sim_mat[i,2]

    #Detection probability
    p1 <- get_p(est[5:6], c(0.0001, est[9]), maxt)
    p2 <- get_p(est[5:6], c(0.0001, est[10]), maxt)
    p3 <- get_p(est[7:8], c(0.0001, est[11]), maxt)

    p1_truth <- get_p(new_truth[5:6], c(0.0001,new_truth[9]), maxt)
    p2_truth <- get_p(new_truth[5:6], c(0.0001,new_truth[10]), maxt)
    p3_truth <- get_p(new_truth[7:8], c(0.0001,new_truth[11]), maxt)
    
    est <- c(est,p1,p2,p3,1/est[5:8])
    new_truth <- c(new_truth, p1_truth, p2_truth, p3_truth, 1/new_truth[5:8])

    data.frame(ID=i, repl=x, param=pars, val=est, truth=new_truth,
               conv=sublist[[x]]$convergence, lik=sublist[[x]]$value)
  })
  do.call("rbind",out)
})
sim_df_N100 <- as_tibble(do.call("rbind", sim_df))

sim_df_N100 <- sim_df_N100 %>%
  left_join(sim_mat, by='ID')

sim_df_N1000 <- lapply(1:9, function(i){
  sublist <- sim_N1000[[i]]
  nsim <- length(sublist)
  out <- lapply(1:nsim, function(x){
    if(is.null(sublist[[x]])) return(data.frame(ID=i, repl=x, param=pars, val=NA, truth=NA,conv=1,
                                     lik=NA))
    est <- sublist[[x]]$par
    psi <- c(exp(est[1]+est[2]+est[3]),exp(est[1]),exp(est[2]),exp(0))
    psi <- psi/sum(psi)
    est <- c(psi, exp(est[4:10]))
    new_truth <- truth
    new_truth[5:11] <- exp(new_truth[5:11])
    new_truth[5] <- sim_mat[i,1]
    new_truth[7] <- sim_mat[i,2]

    #Detection probability
    p1 <- get_p(est[5:6], c(0.0001, est[9]), maxt)
    p2 <- get_p(est[5:6], c(0.0001, est[10]), maxt)
    p3 <- get_p(est[7:8], c(0.0001, est[11]), maxt)

    p1_truth <- get_p(new_truth[5:6], c(0.0001,new_truth[9]), maxt)
    p2_truth <- get_p(new_truth[5:6], c(0.0001,new_truth[10]), maxt)
    p3_truth <- get_p(new_truth[7:8], c(0.0001,new_truth[11]), maxt)
    
    est <- c(est,p1,p2,p3,1/est[5:8])
    new_truth <- c(new_truth, p1_truth, p2_truth, p3_truth, 1/new_truth[5:8])

    data.frame(ID=i+9, repl=x, param=pars, val=est, truth=new_truth,
               conv=sublist[[x]]$convergence, lik=sublist[[x]]$value)
  })
  do.call("rbind",out)
})
sim_df_N1000 <- as_tibble(do.call("rbind",sim_df_N1000))

sim_df_N1000 <- sim_df_N1000 %>%
  left_join(sim_mat, by='ID') %>%
  mutate(ID=ID+9)
```




```{r}

mytheme <- theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
        axis.text=element_text(size=14), axis.title=element_text(size=16),
        legend.text=element_text(size=14), legend.title=element_text(size=16),
        strip.background=element_rect("transparent"), strip.text=element_text(size=16),
        plot.title=element_text(size=16))

occ_sub <- sim_df_N100 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("both","deer","coyote","none")) %>%
  mutate(param=factor(param, levels=c("both","deer","coyote","none")))

occ_sub$padj <- factor(occ_sub$param, levels=c("both","deer","coyote","none"),
                       labels=c("Both~species~(psi[11])","Species~1~only~(psi[10])",
                                "Species~2~only~(psi[\"01\"])","Neither~species~(psi[\"00\"])"))


#Occupancy
occ_sub %>%
  filter(conv==0) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("padj", scales='free', labeller=label_parsed) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  ggtitle("Occupancy, N = 100 sites")
```

**Figure S1**. Proportion bias in model estimates of multi-species occupancy parameters, relative to true values, for a total sample size of N=100 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species. 

\newpage

```{r}
occ_sub <- sim_df_N1000 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("both","deer","coyote","none")) %>%
  mutate(param=factor(param, levels=c("both","deer","coyote","none")))

occ_sub$padj <- factor(occ_sub$param, levels=c("both","deer","coyote","none"),
                       labels=c("Both~species~(psi[11])","Species~1~only~(psi[10])",
                                "Species~2~only~(psi[\"01\"])","Neither~species~(psi[\"00\"])"))


occ_sub %>%
  #filter(conv==0) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("padj", scales='free', labeller=label_parsed) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  ggtitle("Occupancy, N = 1000 sites")
```

**Figure S2**. Proportion bias in model estimates of multi-species occupancy parameters, relative to true values, for a total sample size of N=1000 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species. 

\newpage

```{r}
#Detection
sim_df_N100 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("p1_2pres","p1_2abs","p2")) %>%
  mutate(param=factor(param, levels=c("p1_2pres","p1_2abs","p2")),
         param=fct_recode(param, `Species 1 p | Species 2 present`="p1_2pres", `Species 1 p | Species 2 absent`="p1_2abs",
                          `Species 2 p`="p2")) %>%
  filter(conv==0) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("param", scales='free', nrow=2) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  ggtitle("Detection probability, N = 100 sites")
```

**Figure S3**. Proportion bias in model estimates of overall detection probability, relative to true values, for a total sample size of N=100 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species.

\newpage

```{r}
sim_df_N1000 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("p1_2pres","p1_2abs","p2")) %>%
  mutate(param=factor(param, levels=c("p1_2pres","p1_2abs","p2")),
         param=fct_recode(param, `Species 1 p | Species 2 present`="p1_2pres", `Species 1 p | Species 2 absent`="p1_2abs",
                          `Species 2 p`="p2")) %>%
  filter(conv==0) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("param", scales='free', nrow=2) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  ggtitle("Detection probability, N = 1000 sites")
```

**Figure S4**. Proportion bias in model estimates of overall detection probability, relative to true values, for a total sample size of N=1000 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species.

\newpage

```{r}
#Duration in each state
#sim_df_N100 %>%

#mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
#         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
#  filter(param %in% c("sp1_dur1","sp1_dur2","sp2_dur1","sp2_dur2")) %>%
#  #mutate(param=factor(param, levels=c("p1_2pres","p1_2abs","p2")),
#  #       param=fct_recode(param, `Sp1 p | sp2 present`="p1_2pres", `Sp1 p | sp2 absent`="p1_2abs",
#  #                        `Sp2 p`="p2")) %>%
#  filter(conv==0) %>%
#  mutate(pct_bias = (val-truth)/truth) %>%
#  filter(pct_bias < 10) %>%
#  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
#  geom_boxplot(aes(fill=`Species 2`)) +
#  geom_hline(yintercept=0, linetype=2) +
#  facet_wrap("param", scales='free') +
#  labs(x="Species 1 detectability", y="Proportion bias", fill="Species 2\ndetectability") +
#  mytheme +
#  ggtitle("Duration in each state, N = 100 sites")
```

```{r}
#sim_df_N1000 %>%
#  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
#         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
#  filter(param %in% c("sp1_dur1","sp1_dur2","sp2_dur1","sp2_dur2")) %>%
#  #mutate(param=factor(param, levels=c("p1_2pres","p1_2abs","p2")),
  #       param=fct_recode(param, `Sp1 p | sp2 present`="p1_2pres", `Sp1 p | sp2 absent`="p1_2abs",
  #                        `Sp2 p`="p2")) %>%
#  filter(conv==0) %>%
#  mutate(pct_bias = (val-truth)/truth) %>%
#  filter(pct_bias < 10) %>%
#  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
#  geom_boxplot(aes(fill=`Species 2`)) +
#  geom_hline(yintercept=0, linetype=2) +
#  facet_wrap("param", scales='free') +
#  labs(x="Species 1 detectability", y="Proportion bias", fill="Species 2\ndetectability") +
#  mytheme +
#  ggtitle("Duration in each state, N = 1000 sites")
```

```{r}
#Lambda
sim_df_N100 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("lam1","lam2","lam3")) %>%
  mutate(param=factor(param, levels=c("lam1","lam2","lam3"),
                      labels=c("Sp.~1~lambda~'|'~Sp.~2~present","Sp.~1~lambda~'|'~Sp.~2~absent","Sp.~2~lambda"))) %>%
  filter(conv==0) %>%
  #filter(val < 100) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("param", scales='free', nrow=2, labeller=label_parsed) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  theme(strip.text=element_text(size=14)) +
  ggtitle(expression(lambda*", N = 100 sites"))
```

**Figure S5**. Proportion bias in model estimates of detection intensity ($\lambda$), relative to true values, for a total sample size of N=100 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species.

\newpage

```{r}

sim_df_N1000 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("lam1","lam2","lam3")) %>%
  mutate(param=factor(param, levels=c("lam1","lam2","lam3"),
                      labels=c("Sp.~1~lambda~'|'~Sp.~2~present","Sp.~1~lambda~'|'~Sp.~2~absent","Sp.~2~lambda"))) %>%
  filter(conv==0) %>%
  #filter(val < 100) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("param", scales='free', nrow=2, labeller=label_parsed) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  theme(strip.text=element_text(size=14)) +
  ggtitle(expression(lambda*", N = 1000 sites"))
```

**Figure S6**. Proportion bias in model estimates of detection intensity ($\lambda$), relative to true values, for a total sample size of N=1000 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species.

\newpage

```{r}

#Mu
sim_df_N100 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("mu1[1]","mu1[2]","mu2[1]","mu2[2]")) %>%
  mutate(param=factor(param, levels=c("mu1[1]","mu1[2]","mu2[1]","mu2[2]"),
                      labels=c("Species~1~mu[1]","Species~1~mu[2]","Species~2~mu[1]","Species~2~mu[2]"))) %>%
  filter(conv==0) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  filter(pct_bias < 10) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("param", scales='free', nrow=2, labeller=label_parsed) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  theme(strip.text=element_text(size=14)) +
  ggtitle(expression(mu*", N = 100 sites"))
```

**Figure S7**. Proportion bias in model estimates of the state rate parameters $\mu$, relative to true values, for a total sample size of N=100 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species.

\newpage

```{r}
#Mu
sim_df_N1000 %>%
  mutate(`Species 1`=factor(sp1,levels=c("low","med","high")),
         `Species 2`=factor(sp2,levels=c("low","med","high"))) %>%
  mutate(`Species 1`=fct_recode(`Species 1`, `0.25`="low", `0.50`="med",`1.00`="high")) %>% 
  mutate(`Species 2`=fct_recode(`Species 2`, `0.25`="low", `0.50`="med",`0.68`="high")) %>% 
  filter(param %in% c("mu1[1]","mu1[2]","mu2[1]","mu2[2]")) %>%
  mutate(param=factor(param, levels=c("mu1[1]","mu1[2]","mu2[1]","mu2[2]"),
                      labels=c("Species~1~mu[1]","Species~1~mu[2]","Species~2~mu[1]","Species~2~mu[2]"))) %>%
  filter(conv==0) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  #filter(pct_bias < 10) %>%
  ggplot(aes(y=pct_bias, x=`Species 1`, group=interaction(`Species 1`,`Species 2`))) +
  geom_boxplot(aes(fill=`Species 2`)) +
  geom_hline(yintercept=0, linetype=2) +
  facet_wrap("param", scales='free', nrow=2, labeller=label_parsed) +
  labs(x="Species 1 detection probability", y="Proportion bias", 
       fill="Species 2\ndetection\nprobability") +
  mytheme +
  theme(strip.text=element_text(size=14)) +
  ggtitle(expression(mu*", N = 1000 sites"))
```

**Figure S8**. Proportion bias in model estimates of the state rate parameters $\mu$, relative to true values, for a total sample size of N=1000 sites. Each boxplot represents distribution of bias values for 30 datasets simulated under one of 9 possible scenarios of overall detection probabilities for two species.


```{r}

alldf <- rbind(sim_df_N100, sim_df_N1000) %>%
  mutate(pct_bias = (val-truth)/truth) %>%
  filter(sp1=="high",sp2=="high",N==1000) %>%
  group_by(param) %>%
  summarize(bias_mean=mean(pct_bias),
            bias_sd=sd(pct_bias))

alldf <- alldf[1:10,]

```

