---
title: "Rota_fit_R"
author: "Katie Grabowski"
date: "7/20/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
Collecting and annotating scripts for running occuMulti, an R function that fits Rota's model.

**Might be easier for you to start with the links I sent you, and then look at my thoughts in here? Sorry, it's rather long

my sources: https://rdrr.io/cran/unmarked/man/occuMulti.html and https://cran.r-project.org/web/packages/unmarked/unmarked.pdf

#install package
```{r}
library("unmarked") #install.packages("unmarked")
```

## occuMulti
uses unmarkedFrameOccuMulti to receive data, notes below
```{r}
occuMulti(detformulas, stateformulas, data, maxOrder, starts, method="BFGS",
    se=TRUE, engine=c("C","R"), silent=FALSE, ...)
```
detformulas     Character vector of formulas for the detection models, one per species. 

^^from the pdf: Independent detection models (potentially containing different covariates) must be provided for
each species with the detformulas argument, which takes a character vector of individual formulas with length equal to the number of species s.

#so this would be how we want the detection covariates to be considered? I think these have a bunch of '~1's in them, which means you're suggesting a constant relationship?

stateformulas     Character vector of formulas for the natural parameters. To fix a natural parameter at 0, specify the corresponding formula as "0" or "~0". 

^^from the pdf: The natural parameters can then be modeled as linear functions of covariates. Covariates for each f
must be specified with the stateformulas argument, which takes a character vector of individual formulas of length equal to the number of natural parameters (which in turn depends on the number of species in the model).

#I think natural parameters = f1, f2, f12, etc -- the formulas that include which factors we believe will affect occupancy. which can vary among species

data     An unmarkedFrameOccuMulti object 
#notes below on how to create this

maxOrder     Optional; specify maximum interaction order. Defaults to number of species (all possible interactions). Reducing this value may speed up optimization if you aren’t interested in higher-order interactions. 
#we're only planning to consider pairwise interactions, right? so the maximum interaction order we'd be interested in is 2?

starts     Vector of parameter starting values.
#not sure what this is
#could this be the first replicate survey at each site? that's something the stan code uses

method     Optimization method used by optim.
#should be "BFGS"?
#Method "BFGS" is a quasi-Newton method (also known as a variable metric algorithm), specifically that published simultaneously in 1970 by Broyden, Fletcher, Goldfarb and Shanno. This uses function values and gradients to build up a picture of the surface to be optimized.

se     Logical specifying whether or not to compute standard errors.

engine     Either "C" to use fast C++ code or "R" to use native R code during the optimization.

silent     Boolean; if TRUE, suppress warnings. 
#this sounds rather ominous

#example code for occuMulti
```{r}
## Not run:
#Simulate 3 species data
N <- 1000 #number of cameras? (in other places that's called M)
nspecies <- 3 #often given an 'S', number of species
J <- 5 #max number of sampling days per site

#occ_covs has 1000 rows, 10 columns; I think the 10 is the number of occupancy covariates, and the 1000 might be the number of cameras?
#rnorm generates random deviates (deviate: the value of a random variable measured from some standard point of location, usually the mean)
occ_covs <- as.data.frame(matrix(rnorm(N*10),ncol=10))
names(occ_covs) <- paste('occ_cov',1:10,sep='') #this names the columns
det_covs <- list() #creates an empty list

#N*J is the total number of observations
#for each species, create a matrix with a row for each camera and a column for each day (so 3 matrices in this example)
for (i in 1:nspecies){
det_covs[[i]] <- matrix(rnorm(N*J),nrow=N) 
} #filling in the detection covariates for each species?

names(det_covs) <- paste('det_cov',1:nspecies,sep='') #names the matrices

#True vals
##why is this arbitrarily assigning beta values? 
#10 items, for the 10 occupancy covariates?
beta <- c(0.5,0.2,0.4,0.5,-0.1,-0.3,0.2,0.1,-1,0.1) 

#assuming these are the natural parameter functions, but not sure which relates to which species/species combo
#though that might be answered manually below
#Peter: this is just them providing the "true" values which then they can test the model against (so part of the simulation and not something I need to do to run the model)
f1 <- beta[1] + beta[2]*occ_covs$occ_cov1
f2 <- beta[3] + beta[4]*occ_covs$occ_cov2
f3 <- beta[5] + beta[6]*occ_covs$occ_cov3
f4 <- beta[7]
f5 <- beta[8]
f6 <- beta[9]
f7 <- beta[10]
f <- cbind(f1,f2,f3,f4,f5,f6,f7) #1000 rows, 7 columns
#creates a table with 8 rows (possible species combos, I think--2^3) and 3 columns, one for each species
#only 0 or 1 entries
z <- expand.grid(rep(list(1:0),nspecies))[,nspecies:1]
colnames(z) <- paste('sp',1:nspecies,sep='') #names the columns in z

#paste0()converts its arguments to character strings and concatenates them
#this creates a spreadsheet with columns for sp 1, sp 2, sp 3, sp 1:sp 2, sp 1:sp 3, sp 2:sp 3, sp 1:sp 2:sp 3, which seems promising? 7 columns (for each f) and 8 rows (for each spp combo)
#all entries are 1 or 0
#model.matrix creates a design (or model) matrix, e.g., by expanding factors to a set of dummy variables (depending on the contrasts) and expanding interactions similarly.
#this is the design matrix?!
#Peter: rows are all species presence/absence combinations, columns are: each species(3), each possible pair(3), and all 3; with a 1 in the cell if all the species in the corresponding species combination are included in the column header --> a truth table
dm <- model.matrix(as.formula(paste0("~.^",nspecies,"-1")),z)

#matrix multiplication; exponential function of (f times the transpose of the design matrix)
psi <- exp(f %*% t(dm)) #1000 rows x 8 columns
psi <- psi/rowSums(psi) #Peter:probability of each different species combo (because we're simulating data, we just make that up)

#True state
#makes an empty matrix
#Peter: ztruth is just being used to create the fake detection histories; math to make a detection history for each species that matches your proposed psi
ztruth <- matrix(NA,nrow=N,ncol=nspecies) #1000 rows, 3 columns

#fills the matrix but not clear to me how/why
#what does ztruth represent?
for (i in 1:N){
ztruth[i,] <- as.matrix(z[sample(8,1,prob=psi[i,]),])
}
p_true <- c(0.6,0.7,0.5) #huh?
# fake y data
##I think these are detection histories for each spp (so then the rbinom function is just making up whether it should be 1 or 0 because this isn't real data)
y <- list()
for (i in 1:nspecies){ #for each species
y[[i]] <- matrix(NA,N,J) #blank matrix, 1000 x 5
for (j in 1:N){ #for each camera
  for (k in 1:J){ #for each day
y[[i]][j,k] <- rbinom(1,1,ztruth[j,i]*p_true[i]) #Peter: coming up with the 1/0 for each species at each site on each day (j,i:each camera, each species); p_true is weighting it by something else, maybe by detection probabilities)
}
}
}
names(y) <- c('coyote','tiger','bear') #y contains 3 matrices

#Create the unmarked data object
##wait...are the detection covariates the observation covariates here? I thought observation covariates varied within a site (this is also kind of where I lost what was going on)
data = unmarkedFrameOccuMulti(y=y,siteCovs=occ_covs,obsCovs=det_covs)

#Summary of data object
summary(data)
plot(data)
# Look at f parameter design matrix
data@fDesign

# Formulas for state and detection processes
# Length should match number/order of columns in fDesign
#Peter: this is saying that the first species is a function only of the first occupancy covariate; should be able to add other covariates (maybe with a +)
#Peter: trying to figure out what the '~1' means; seems to be that it's constant (not dependent on any of the covariates)
#Peter: says that including lat/long/combo might not be necessary for our size site
#Peter: putting something in here for the columns representing species pairs is looking at the effect of that covariate on the probability of seeing those two species in the same place
occFormulas <- c('~occ_cov1','~occ_cov2','~occ_cov3','~1','~1','~1','~1')
#Length should match number/order of species in data@ylist
#Peter: whatever the detection covariates are modeling is a constant
detFormulas <- c('~1','~1','~1')
fit <- occuMulti(detFormulas,occFormulas,data)

#Look at output
fit
plot(fit)
#Compare with known values
#Peter: looking at what we made up vs what the model estimated
cbind(c(beta,log(p_true/(1-p_true))),fit@opt$par)
#predict method
#Peter: predict gives the model estimates for a given thing, like state (probabilities of the different species combos) or det
#predict gives a predicted value for each camera, unsure what it is? potentially when you call for a certain species, the likelihood of seeing that species at the camera
lapply(predict(fit,'state'),head)
lapply(predict(fit,'det'),head)
#marginal occupancy
head(predict(fit,'state',species=2))
head(predict(fit,'state',species='bear'))
head(predict(fit,'det',species='coyote'))
#probability of co-occurrence of two or more species
head(predict(fit, 'state', species=c('coyote','tiger')))
#conditional occupancy
head(predict(fit,'state',species=2,cond=3)) #tiger | bear present
head(predict(fit,'state',species='tiger',cond='bear')) #tiger | bear present
head(predict(fit,'state',species='tiger',cond='-bear')) #bear absent
head(predict(fit,'state',species='tiger',cond=c('coyote','-bear')))

#residuals (by species)
#Peter: residual is how wrong the model is, smaller is better
lapply(residuals(fit),head)

#Peter: not sure what ranef is
#ranef (by species)
ranef(fit, species='coyote')
#parametric bootstrap
bt <- parboot(fit,nsim=30)

#update model
occFormulas <- c('~occ_cov1','~occ_cov2','~occ_cov2+occ_cov3','~1','~1','~1','~1')
fit2 <- update(fit,stateformulas=occFormulas)

#List of fitted models
#Peter: used to compare the two models
fl <- fitList(fit,fit2)
coef(fl)

#Peter: this pulls out different ways of comparing the models
#Model selection
modSel(fl)
#Fit model while forcing some natural parameters to be 0
#For example: fit model with no species interactions
occFormulas <- c('~occ_cov1','~occ_cov2','~occ_cov2+occ_cov3','0','0','0','0')
fit3 <- occuMulti(detFormulas,occFormulas,data)
#Alternatively, you can force all interaction parameters above a certain order to be zero with maxOrder. This will be faster.
occFormulas <- c('~occ_cov1','~occ_cov2','~occ_cov2+occ_cov3')
fit4 <- occuMulti(detFormulas,occFormulas,data,maxOrder=1)
## End(Not run)
```


## unmarkedFrameOccuMulti
```{r}
unmarkedFrameOccuMulti(y, siteCovs=NULL, obsCovs=NULL,
maxOrder, mapInfo)

```
arguments:
y     A list (optionally a named list) of length S where each element is an MxJ matrix of the detection, non-detection data for one species, where M is the number of sites, J is the maximum number of sampling periods per site, and S is the number of species in the analysis. 
#this corresponds to our detection histories

siteCovs     A data.frame of covariates that vary at the site level. This should have M rows and one column per covariate 
#cool, this is the psi and p covariate sheets, not sure how you tell it which one is for occupancy vs detection probabilities (unless it's only the occupancy-related covariates, and there's another way of incorporating the detection ones?)

obsCovs     Either a named list of data.frames of covariates that vary within sites, or a data.frame with MxJ rows in site-major order. 
#no idea. what type of covariate would vary within a site? I think this would be NULL for our data. 
#Unless these are the detection covariates? some hint of that, maybe, above (where it names obsCovs = det_covs when running occuMulti)

maxOrder     Optional; specify maximum interaction order. Defaults to number of species (all possible interactions). Reducing this value may speed up creation of unmarked frame if you aren’t interested in higher-order interactions.
#comment above 

mapInfo     Currently ignored

#example code for unmarkedFramOccuMulti
```{r }
# Fake data
S <- 3 # number of species
M <- 4 # number of sites
J <- 3 # number of visits

#is the below code just creating a matrix for use? with made up data
y <- list(matrix(rbinom(M*J,1,0.5),M,J), # species 1
matrix(rbinom(M*J,1,0.5),M,J), # species 2
matrix(rbinom(M*J,1,0.2),M,J)) # species 3
site.covs <- data.frame(x1=1:4, x2=factor(c('A','B','A','B')))
site.covs #4 rows (number of sites) x 2 columns (2 covariates?)
umf <- unmarkedFrameOccuMulti(y=y, siteCovs=site.covs,
obsCovs=NULL) # organize data
umf # look at data
summary(umf) # summarize
plot(umf) # visualize
#fm <- occu(~1 ~1, umf) # fit a model

```
