---
title: "multi-season sample"
author: "Katie Grabowski"
date: "11/20/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

#install packages (starting with these because they seem like a good group)
```{r}
library("unmarked")  #install.packages("unmarked") 
library(camtrapR) #install.packages("camtrapR")
library(tidyverse)
library(dplyr)
library(patchwork) #install.packages("patchwork")
library(ggpubr)  #install.packages("ggpubr")
```

## get sample data
We have three covariates that vary by site: median elevation of the quadrat (ele, in metres), forest cover of the quadrat (forest, in percent) and the number of surveys per season (i.e., 2 or 3, surveys). These are called site covariates, because they vary by sites only. The 27 columns entitled "det991" to "det073" contain the crossbill detection/nondetection data during all surveys over the 9 years. They contain a 1 when at least one crossbill was recorded during a survey and a 0 otherwise. NAs indicate surveys that did not take place, either because a site is high-elevation and has no third survey or because it failed to be surveyed altogether in a year. The final 27 columns entitled "date991" to "date073" give the Julian date of each survey. They represent a `survey-covariate' or `observation covariate'. We note that the paper by Royle and Kery (2007) used a subset of this data set.

summary: 
each row corresponds to a site
each site has 3 site covariates (vary by site only): elevation, forest cover and number of surveys (1, 2 or 3)
the next 27 columns give detection/non-detection data (0 or 1) for all of the surveys, with NAs if a survey wasn't conducted
the final 27 columns give the Julian data of each survey; these are a 'survey-covariate' or 'observation-covariate'


```{r}
#get data from unmarked package
data(crossbill)
colnames(crossbill)
```

## Adjusting dates

unmarked removes missing data in a context specific way. For missing siteCovs, the
entire row of data must be removed. (so if you don't have a base feature of the site, like elevation or forest cover here, the whole site is scrapped)

However, for missing yearlySiteCovs or obsCovs, only
the corresponding observation are removed. 

Thus, if unmarked removes different
observations from different models, the models cannot be compared using AIC. A way
around this is to remove the detection data corresponding to missing covariates before
fitting the models.

Notes: 
is.na returns FALSE if there's a value and TRUE if NA
```{r}
##The crossbill data have missing dates and so we remove the associated detection/non-detection data.
DATE <- as.matrix(crossbill[,32:58]) #select columns with the Julian date of each survey
y.cross <- as.matrix(crossbill[,5:31]) #select crossbill detection/nondetection data during all surveys over the 9 years
y.cross[is.na(DATE) != is.na(y.cross)] <- NA #I think this sets any values where either the date or the detection data is missing equal to NA (if both have data, nothing happens, and if neither has data, it's already NA, I believe)
```

In addition, continuous covariates should be transformed in a way that brings their
values close to zero in order to improve or even enable numerical convergence of the
maximum-likelihood routine. 

```{r}
##scale data
DATE <- scale(DATE)
```

#Format data appropriately

Notes:
y.cross has a row for every site and detection/non-detection data for every survey across 9 years 

primary sample period: year or season
secondary sample period: number of surveys/days/observations per primary sample period

```{r}
years <- as.character(1999:2007) #I believe this just creates a list with the relevant years? another example of using colext does the same
years <- matrix(years, nrow(crossbill), 9, byrow=TRUE) #creates a matrix with each of the years across the rows, and a row for each site (another colext examples do the same)
umf <- unmarkedMultFrame(y=y.cross, #creates the actual data object; sets y to y.cross (matrix of observed data)
siteCovs=crossbill[,2:3], yearlySiteCovs=list(year=years), #assigns siteCovs to the second and third columns of crossbill (elevation and forest); assigns the list of years as the yearlySiteCovs (covariates at the site-year level)
obsCovs=list(date=DATE), #sets obsCovs (covariates that vary within site-year-observation level) to DATE
numPrimary=9) #number of primary time periods (in this case, years)
```

#Model fitting
We fit a series of models that represent different hypotheses about the
colonization-extinction dynamics of Swiss crossbills at a spatial scale of 1 km2. We fit year
effects on colonization and extinction in the means parameterization, but for detection
probability, we choose an effects parameterization. The latter is more useful for getting
predictions in the presence of other explanatory variables for that parameter. For model
fm5 with more complex covariate relationships, we use as starting values for the
optimization routine the solution from a "neighboring" model with slightly less complexity,
model fm4. Wise choice of starting values can be decisive for success or failure of maximum
likelihood estimation.

colext(psiformula= ~1, gammaformula =  ~ 1, epsilonformula = ~ 1,
    pformula = ~ 1, data, starts, method="BFGS", se=TRUE, ...)

psiformula: right-hand sided formula for the initial probability of occupancy at each site
gammaformula: right-hand sided formula for colonization probability
epsilonformula: right-hand sided formula for extinction probability
pformula: right-hand sided formula for detection probability
data: unmarkedMultFrame object that supplies the data
starts: optionally, initial values for parameters in the optimization
method: optimization method used by optim
se: logical specifying whether or not to compute standard errors

```{r}
 # A model with constant parameters
 fm0 <- colext(~1, ~1, ~1, ~1, umf)

# Look at output
fm0

 # Like fm0, but with year-dependent detection
 fm1 <- colext(~1, ~1, ~1, ~year, umf)
 
 # Like fm0, but with year-dependent colonization and extinction
 fm2 <- colext(~1, ~year-1, ~year-1, ~1, umf)
 
 # A fully time-dependent model
 fm3 <- colext(~1, ~year-1, ~year-1, ~year, umf)
 
 # Like fm3 with forest-dependence of 1st-year occupancy
 fm4 <- colext(~forest, ~year-1, ~year-1, ~year, umf)
 
 # Like fm4 with date- and year-dependence of detection
 fm5 <- colext(~forest, ~year-1, ~year-1, ~year + date + I(date^2),
umf, starts=c(coef(fm4), 0, 0))
 
 # Same as fm5, but with detection in addition depending on forest cover
 fm6 <- colext(~forest, ~year-1, ~year-1, ~year + date + I(date^2) +
forest, umf)

```

#Model selection
Use AIC to compare models

Model selection and model-averaged prediction in unmarked require that we create a
list of models using fitList. This function organizes models and conducts a series of tests to
ensure that the models were fit to the same data.

```{r}
#Create an unmarkedFitList [this method is unnecessarily complicated]
#models <- fitList('psi(.)gam(.)eps(.)p(.)' = fm0,
#'psi(.)gam(.)eps(.)p(Y)' = fm1,
#'psi(.)gam(Y)eps(Y)p(.)' = fm2,
#'psi(.)gam(Y)eps(Y)p(Y)' = fm3,
#'psi(F)gam(Y)eps(Y)p(Y)' = fm4,
#'psi(F)gam(Y)eps(Y)p(YD2)' = fm5,
#'psi(F)gam(Y)eps(Y)p(YD2F)' = fm6)

#Model selection results from an unmarkedFitList
#ms <- modSel(models)

#view output
#ms

#Why can'tI just do this to create a fitlist? Answer: I can
fl <- fitList(fm0, fm1, fm2, fm3, fm4, fm5, fm6)

#Model selection results from an unmarkedFitList
ms <- modSel(fl)

#View output
ms

coef(ms) # Estimates only
SE(ms) # Standard errors only
toExport <- as(ms, "data.frame") # Everything
```

#Prediction and plotting
Stopped here until I better understand what's going on until this