---
title: "Gorongosa_occupancy"
author: "Kaitlyn Gaynor & Lindsey Rich"
date: "2/28/20"
output: html_document
editor_options: 
  chunk_output_type: console
---

Multi-species occupancy modeling

Get set-up. Load libraries, etc.
```{r setup}
library(R2jags)
library(reshape)
library(reshape2)
library(plyr)
library(dplyr)
library(here)
library(tidyverse)
```

Define function for standardizing the covariates. You'll need this later.
```{r stdev}
zstd <- function (x)  (x-mean(x,na.rm = T))/sd(x,na.rm = T)  
```

# Import and manipulate detection/operation histories.

```{r import seasonal detections - rainfall}
D <- read.csv("scripts/alternative-multispecies-model/detections-dry2016_katie-8_1_16-11_30_16.csv")
Camop <- read.csv("scripts/alternative-multispecies-model/camoperation-dry2016_katie-8_1_16-11_30_16.csv")
```


# Import the rest of the data (species ID/traits, metadata).
```{r import other data}
Spp <- read.csv("~/Documents/github-repos/gorongosa/gorongosa-camera-traps/data/2018spp_kingdon.csv") # Species IDs and traits
Covariates <- read.csv("data/gorongosa-cameras/cam_metadata_fromfield_and_raw_raster.csv")

# Reorganize detection dataframe, so that it has one column for species and one for detections (rather than a column for each species)
D <- melt(D,vars = c("StudySite"))
colnames(D)[2:3] = c("SppCode","Detections")
head(D) # make sure it worked

# Combine detection history with StudySite covariates and species traits
D <- merge(D, Covariates, by = "StudySite") # joins by StudySite
D <- merge(D, Camop, by = "StudySite") # joins by StudySite
D <- merge(D, Spp, by = "SppCode") # joins by SppCode
head(D)

# filter out only carnivores (but not lions)
D <- D %>% filter(Order == "Carnivora") %>% filter(SppCode != "PALE")

# drop levels
D$SppCode <- droplevels(D$SppCode)

# sort to get back in order
D <- arrange(D, SppCode, StudySite)
```

Save species names for use later when exporting model results
```{r save species names}
# sppnames2 is just missing 'SppCode' line
sppnames <- c("SppCode", as.character(unique(D$SppCode)))
sppnames2 <- as.character(unique(D$SppCode))
```

# Run model

This model includes all six covariates from original analysis.
```{r dry season both years with covariates}
# select occupancy covariates
X = dplyr::select(D, termite.count.100m, urema_dist, tree_hansen)
# normalize them
X = apply(X,2,zstd)

## select detection covariates
dX = dplyr::select(D, detect.obscured, cover.ground)

# normalize them
dX = apply(dX,2,zstd)

# Define data and parameters. Load all the data including the detection array, number of sampling occasions, individual species sampled, total number of sampled species, and covariate information.
data <- list(D = D$Detections, # number of detections
             N = ceiling(D[,"Operation"]), # number of trap-nights for each StudySite
             Species = as.numeric(as.factor(D$SppCode)), # species
             n = nrow(D), 
             nspp = max(as.numeric(as.factor(D$SppCode))),
             X = X, 
             dX = dX)

# specify the initial values
inits = function() {list(Z = as.numeric(data$D>0))}

# specify the parameters to be monitored
params = c("rho","pbeta","spbeta","sigpbeta","mbeta","sigbeta","sbeta","psi.mean","sigma.occ","p.mean","sigma.p","alpha","Z","P")

nc = 3       # number of chains
ni = 60000   # number of iterations
nb = 10000  # burn-in period
nthin = 50   # thinning rate


# Run occupancy model. The model file must be in the correct folder (see script AllMammals_pcov.txt)
out3 <- jags(data = data, 
             inits = inits, 
             parameters.to.save = params, 
             model.file ="scripts/alternative-multispecies-model/AllMammals_pcov_nogroup.txt", 
             n.chains = nc, 
             n.iter = ni,
             n.burnin = nb, 
             n.thin = nthin)

out3.sum <- out3$BUGSoutput$summary
```

# Export the results

### Occupancy and detection probabilities
```{r}
write.table(x=out3.sum,file="scripts/alternative-multispecies-model/results/mesocarnivore.csv",sep=",") # save output

#Name this output alpha and P.
alpha <- out3$BUGSoutput$sims.list$alpha
p <- out3$BUGSoutput$sims.list$P

# we are log transforming alphas b/c they are on logit scale.
expit <- function(x)  1/(1+exp(-x))
logit.alpha <- expit(alpha)
logit.p <- expit(p)

# Columns represent species, so we take the mean of each column to get the mean psi and p value (when covariates are at their mean values) for each species.

psimeans <- colMeans(logit.alpha)
names(psimeans) <- sppnames2
psimeans <- as.data.frame(psimeans)
write.table(x = psimeans, file = "scripts/alternative-multispecies-model/results/mesocarnivore_alphaspsi()p().csv", sep = ",")

pmeans <- colMeans(logit.p)
names(pmeans) <- sppnames2
write.table(x = pmeans, file = "scripts/alternative-multispecies-model/results/mesocarnivore_detection.csv", sep=",")

# Get the quantiles and 95% confidence intervals for psi and p.

apply(logit.alpha, 2, function(x) sort(x)[])
psiCI <- apply(logit.alpha, 2, function(x) quantile(x,probs = c(0.025,0.1,0.5,0.9,0.975)))
colnames(psiCI) <- sppnames2
write.table(x=psiCI, file="scripts/alternative-multispecies-model/results/mesocarnivore_alphaCI.psi()p().csv",sep=",")

apply(logit.p, 2, function(x) sort(x)[])
pCI <- apply(logit.p, 2, function(x) quantile(x,probs = c(0.025,0.1,0.5,0.9,0.975)))
colnames(pCI) <- sppnames2
write.table(x = pCI, file="scripts/alternative-multispecies-model/results/mesocarnivore_pCI.psi()p().csv", sep = ",")
```

### Occupancy covariate effects
```{r}
# Define the occupancy covariate effects where mbeta is the community-level hyperparameter and sbeta is the species-specific parameter.
mbeta <- out3$BUGSoutput$sims.list$mbeta
sbeta <- out3$BUGSoutput$sims.list$sbeta

# Define the species
spec <- unique(D$SppCode)

# Define the occupancy covariates
covs <- colnames(X)

# Create a data frame where the number of rows is equal to the number of covariates * the number of species
species <- data.frame(expand.grid(covs,spec), matrix(NA,length(covs)*length(spec),4))
colnames(species) <- c("Factor","Species","Mean","SD","LCI","UCI")

# Create a loop that will estimate species-specific values for each of the covariates
for (a in 1:length(covs)){
  for (b in 1:length(spec)){
    sims <- mbeta[,a] + sbeta[,b,a]
    species[(ncol(X)*(b-1)+a),3:6] <- c(mean(sims),sd(sims),quantile(sims,c(0.025,0.975)))
  }
}

# export the table
write.table(x=species,file="scripts/alternative-multispecies-model/results/mesocarnivore_species.csv",sep=",", row.names=F) 

# calculate distributions for community hyperparameters
comm_param <- as.data.frame(mbeta)
names(comm_param) <- covs
head(comm_param)
comm_param_long <- comm_param %>% 
  pivot_longer(cols = everything(), names_to = "Factor", values_to = "Value")
write.csv(comm_param_long, "scripts/alternative-multispecies-model/results/mesocarnivore_communityposteriors.csv", row.names = F)
```

## Species richness
```{r}
# Species richness for each site

# Define the z matrix
z = out3$BUGSoutput$sims.list$Z

# Sort the data frame based on species, study site
d <- sort_df(merge(data.frame(ID = 1:nrow(D),D[,1:2]),data.frame(SppCode = spec)),"ID")[,c(1,3)]

# Create a new data frame
dz <- data.frame(d,t(z))

# Melt the data frame for easy casting
m.dz <- melt(dz,id.vars = c("SppCode","StudySite"))

# Aggregate the data by summing the values in the z matrix for each camera station during each iteration
z.all <- acast(m.dz,StudySite ~ variable, fun.aggregate = sum)

write.table(x=z.all,file="scripts/alternative-multispecies-model/results/mesocarnivore_rawrichnessdistribution.csv",sep=",")

# Use the aggregated values to create probability distributions and estimate mean, sd, and 95% credible interval values for camera-station specific species richness
z.all <- t(apply(z.all,1,function(x) c(mean(x),sd(x),quantile(x,c(0.025,0.975)))))
names <- rownames(z.all)
rownames(z.all) <- NULL
z.all <- cbind(names,z.all)
colnames(z.all) = c("StudySite", "Mean","SD","LCI","UCI")
# not working anymore, hmm, skipping for now

# Export estimates of species richness as a table
write.table(x=z.all,file="scripts/alternative-multispecies-model/results/mesocarnivore_spprich.csv",sep=",", row.names=F)
```
